Temporality is Important for Networks of Hospital Patient Transfers

### 2024-10-30 reports/report.qmd
Add report content and diagram filter

---
title: "Temporality is Important for Networks of Hospital Patient Transfers"
authors:
    - name: David Wu
    - name: Michael Lydeamore
    - name: Andrew Stewardson
    - name: Tjibbe Donker
    - name: Nic Geard
    - name: "******"
date: last-modified
format:
    html: default
    pdf: default
bibliography: refs.bib
csl: plos.csl
---

@@

Understanding of the spread of AMR through hospital systems has been studied in the past as a network phenomenon. The movement of patients through the hospital systems induced by referrals can be represented as edges between hospital nodes. However, these networks are usually presented as static, where movement over a large period of time has been aggregated to form a single network. Typically, this is due to the lack of granularity in the base dataset. It is known, especially in other applications, that temporal networks can display different behaviour than the static equivalents. In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison.

@@

## Network Modelling

@@

@@

We simulate the movement of individuals

@@

>>>>

---
title: "Temporality is Important for Networks of Hospital Patient Transfers"
authors:
    - name: David Wu
    - name: Michael Lydeamore
    - name: Andrew Stewardson
    - name: Tjibbe Donker
    - name: Nic Geard
    - name: "******"
date: last-modified
format:
    html: default
    pdf: default
bibliography: refs.bib
csl: plos.csl
jupyter: python3
filters:
    - diagram
execute:
    echo: false
    warning: false
    error: false
---

```{python}
import numpy as np
import polars as pl
from matplotlib import pyplot as plt
import seaborn as sns
import itertools
import re

from pathlib import Path

# setup printing formats for tables
pl.Config.set_tbl_hide_column_data_types(True)
pl.Config.set_tbl_hide_dataframe_shape(True)

sns.set_theme(style='ticks')
sns.set_context('paper')
# plt.style.use('ggplot')
```

@@

Understanding of the spread of AMR through hospital systems has been studied in the past as a network phenomenon. The movement of patients through the hospital systems induced by referrals can be represented as edges between hospital nodes. However, these networks are usually presented as static, where movement over a large period of time has been aggregated to form a single network. It is known, especially in other applications, that temporal networks can display different behaviour than the static equivalents. In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison. In particular, we investigate the impacts of the inclusion of modelling the period of time that patients spend at home between admissions, and the effect of approximating the granularity of the temporality]]]]].

@@

## Network Modelling {#sec-network-modelling}

@@

```{python}
#| label: fig-readmission-window
#| fig-cap: Distribution of readmission time (time between discharge and successive admission of a single patient)

# sourced from readmission_time.py
readmit_window = pl.read_parquet("../concordant_networks/readmission_time.parquet")

fig, ax = plt.subplots()
ax.scatter(*(readmit_window.to_numpy().T), marker='.')
ax.set_xscale('asinh') # todo fix xticks
ax.set_yscale('log')
ax.set_xlabel('Readmission Window (days)')
ax.set_ylabel('Count')
ax.set_aspect(0.5)
ax.set_xticks([0, 1, 1e1, 1e2])
ax.set_xticks([2, 5, 2e1, 5e1, 2e2, 5e2], minor=True)
sns.despine()
```

{skip}

### Networks and Projections

Patient transfers can be considered as edges from some source hospital $x_i$ at some time $t_m$ to a target hospital $x_j$ at some time $t_n$. We could work with the exact times, but this will lead to an overly sparse representation that is difficult to work with.
Instead, we discretise time into bins by treating movements involving some arbitrary hospital $x_i$ between times $t_n$ and $t_n + W$ as instead involving a single node at $v_{i,n} \coloneqq (x_i, t_n)$. 
Then we can have a temporal network that can be treated as a static network as in @sec-network-modelling, i.e. we have edge weights $w_{ij}$ along the edge $e_{ij} \coloneqq (v_i, v_j)$.

The network analysis of the temporal network as described above is computationally expensive since it is relatively large ($N_x = 388$ hospitals and $N_t = 259$ fortnightly bins, for $\approx 10000$ nodes).
We can instead analyse slices of the temporal network, where a slice at $t_n$ consists of nodes at that time discretisation bin, and the related nodes.

### Dynamics

<!-- TODO need to work on indexing consistency -->

For the temporal network, we can model the number of patients discharged from some given location ($x_i$) at a given time ($t_n$) as a Poisson random variable with rate $\gamma(x_i, t_n)$. Some proportion $q(x_i, \emptyset)$ of these individuals will not re-admit at any other hospital, and can be removed from the system; the remaining discharged patients will have some probability $p(x_i, t_n, x_j, t_m)$ to re-admit at location $x_j$ at time $t_m$. These probabilites $p$ wil depend on the edges...
Patients that have $t_i \neq t_j$ will be moved into an auxiliary state $z_{i,j}$...
Movemnts $z_{ij}$ are distriuted as Uniform(0, W)

@@

@@

```{python}
# simulation archives from run_zero*.py
simulation_archives = {
    'temporal': '../zero_sims/temporal/sim_all_30s.h5',
    'static': '../zero_sims/static/sim_all_30s.h5',
    'snapshot': '../zero_sims/snapshot/sims_with_30seeds_pseudocap.h5',
    'naive_static': '../zero_sims/naive_static/sim_all_30s.h5',
}

processed_results = {}

# metrics precomputed in compute_summary_stats.ipynb
for label, archive in simulation_archives.items():
        processed_results[label] = pl.read_parquet(Path(archive).with_name('metrics_30s.parquet'))
```

We define the _extent_ of a simulation by the number of facilities that some patient has moved to or through within a given number of days. We see in @fig-extent-30 that we get starkly different distriubtions of extent for the naïve static model, commpared to the other models.

```{python}
#| label: fig-extent-30
#| fig-cap: "Distribution of 30-day extent in different simulation models on different network types"

combined_all_df = pl.concat(
    [(df.select('seed', 'extent')
        .with_columns(
            pl.lit(label).alias('simulation'), 
            pl.col('extent').cast(pl.Int64)
        )
     ) for label, df in processed_results.items()
    ]
)

sim_type_order = ['temporal', 'snapshot', 'static', 'naive_static']
extent_bins = np.arange(0, combined_all_df.select('extent').max().item()+1)

sns.histplot(
    combined_all_df, 
    x = 'extent', 
    hue = 'simulation', 
    multiple = 'dodge', 
    bins = extent_bins, 
    shrink = 0.8, 
    hue_order = sim_type_order,
)
sns.despine()
```

```{python}
# comparisons from compute_summary_stats.ipynb
comparative_results = dict()
name_pattern = re.compile(r"(\w+)_x_([\w_]+)\.30s")
for comp_file in Path("../zero_sims/comparisons/").glob("*.30s.parquet"):
    pair = name_pattern.match(comp_file.stem).groups()
    if pair is None:
        raise RuntimeError(f"Unable to parse file name: {comp_file.name}")
    paired_comparison = pl.read_parquet(comp_file)
    comparative_results[pair] = paired_comparison
    comparative_results[tuple(reversed(pair))] = paired_comparison.with_columns(-pl.col('statistic'))
```

```{python}
#| label: tbl-example
#| tbl-cap: An example table

comparative_results['static', 'snapshot'].head(5)
```
